How to obtain the redshift distribution from probabilistic redshift estimates

### 2019-05-20 research/paper/draft.tex
updating files

@@

\usepackage{tikz}
\usepackage{natbib}
\usepackage{amsmath}

@@

\usepackage{soul}
\usepackage[title]{appendix}

\newcommand{\todo}[3]{{\color{#2}\emph{#1}: #3}}
\newcommand{\aim}[1]{\todo{AIM}{red}{#1}}
\newcommand{\myemail}{aimalz@nyu.edu}
\newcommand{\mathul}[1]{\underline{#1}}
\newcommand{\Sect}[1]{Section~\ref{#1}}
\newcommand{\Eq}[1]{Equation~\ref{#1}}
\newcommand{\Fig}[1]{Figure~\ref{#1}}
\newcommand{\project}[1]{{\textsc{#1}}~}
\newcommand{\lsst}{\project{LSST}}
\newcommand{\desc}{\lsst-\project{DESC}}
\newcommand{\sdss}{\project{SDSS}}
\newcommand{\boss}{\project{BOSS}}
\newcommand{\des}{\project{DES}}
\newcommand{\Chippr}{\project{CHIPPR}}
\newcommand{\repo}[1]{{\texttt{#1}}~}
\newcommand{\qp}{\repo{qp}}
\newcommand{\chippr}{\repo{chippr}}
\newcommand{\cosmolike}{\repo{CosmoLike}}
\newcommand{\emcee}{\repo{emcee}}
\newcommand{\github}{\href{https://github.com}{GitHub}}
\newcommand{\python}{\textit{Python}}
\newcommand{\nz}{$n(z)$}
\newcommand{\data}{\ensuremath{\vec{d}}}
\newcommand{\like}{\mathscr{L}}
\newcommand{\pr}[1]{\ensuremath{\mathrm{p}(#1)}}
\newcommand{\expect}[1]{\left<#1\right>}
\newcommand{\normal}[2]{\mathcal{N} (#1, #2)}
\newcommand{\gvn}{\mid}% could use | or \vert
\newcommand{\integral}[2]{\ensuremath{\int\ #1\ \mathrm{d} #2}}
\newcommand{\sz}{spec-$z$~}
\newcommand{\Sz}{Spec-$z$~}
\newcommand{\pz}{photo-$z$~}
\newcommand{\Pz}{Photo-$z$~}
\newcommand{\pzpdf}{\pz PDF}
\newcommand{\Pzpdf}{\Pz PDF}
\newcommand{\zpdf}{redshift posterior}
\newcommand{\pzip}{\pz implicit posterior}
\newcommand{\Nz}{$N(z)$~}
\newcommand{\stack}{$\hat{N}(z)$~}
\newcommand{\ntot}{\ensuremath{N_{\mathrm{tot}}}}
\newcommand{\bvec}[1]{{\ensuremath{\boldsymbol{#1}}}}
\newcommand{\ndphi}{\bvec{\phi}}
\newcommand{\mmle}{marginalized maximum likelihood estimate}

\begin{document}

@@

\author{Alex Malz\altaffilmark{1}}
\author{David W. Hogg\altaffilmark{1,2,3,4}}
\email{aimalz@nyu.edu}

@@

\begin{abstract}
A trustworthy estimate of the redshift distribution \nz\ is crucial for using 
galaxy catalogs to study cosmology via weak gravitational lensing and 
large-scale structure.
Spectroscopic redshifts for the dim and numerous galaxies of weak-lensing 
surveys are expected to be inaccessible, making photometric redshifts 
(photo-$z$s) the next-best alternative.
The nontrivial systematics affecting photo-$z$ estimation have motivated the 
weak-lensing community to favor photo-$z$ probability density functions (PDFs) 
as a more comprehensive alternative to photo-$z$ point estimates.
However, analytic methods for utilizing these new data products in cosmological 
inference are still evolving.
The ubiquitous methodology known as stacking produces a systematically biased 
estimator of $n(z)$ that worsens with decreasing signal-to-noise, the very 
regime where photo-$z$ PDFs are most necessary.
We introduce a mathematically rigorous probabilistic graphical model (PGM) of 
hierarchical inference of $n(z)$, which is provably the only self-consistent 
way to combine photo-$z$ PDFs to produce an estimator of $n(z)$.
The novel Cosmological Hierarchical Inference with Probabilistic Photometric 
Redshifts (\Chippr) model yields a more accurate characterization of $n(z)$ by 
correctly propagating the redshift uncertainty information beyond the best-fit 
estimator produced by traditional procedures.
We conclude by propagating these effects to constraints in the space of 
cosmological parameters.
\end{abstract}

@@

@@

\aim{TODO: Pull in content from global intro to thesis.}

Though their potential to improve estimates of physical parameters is 
tremendous, \pzpdf s have been applied only to a limited extent.  
They have been used to form selection criteria of samples from galaxy surveys 
without propagation 
through the calculations of physical parameters 
\citep{VanBreukelen2009,Viironen2015}.  
Probability cuts on Bayesian quantities are not uncommon \citep{Leung2015, 
DiPompeo2015a}, but that procedure does not fully take advantage of all 
information contained in a probability distribution for parameter inference.

@@

The \textit{redshift distribution function \Nz}, or, almost interchangably, its 
normalized cousin the \textit{redshift density function \nz}, serves as an 
ideal statistic upon which to demonstrate this novel approach.  
\Nz\ is necessary for calculations of two-point correlation functions of weak 
gravitational lensing and counting statistics that are used to probe dark 
energy \citep{Masters2015}.  
The observed \Nz\ also been used to validate survey selection functions 
usnfenjed in generation of realistic, multi-purpose mock catalogs 
\citep{Norberg2002}.  
Additionally, \Nz\ has been the subject of inference using \pzpdf s before 
\citep{sheldon_photometric_2012, Hildebrandt2012, Kelly2014, Benjamin2013, 
bonnett_using_2015, Viironen2015, Asorey2016, leistedt_hierarchical_2016}, so 
comparisons to the literature may easily be made.

The prevailing way to estimate \nz\ for a sample of $\ntot$ galaxies $i$ is to 
``stack'' \pzpdf s according to Equation~\ref{intro:eq:eqn:stack}.
The stacked estimator of the redshift density function is effectively an 
average of the \pzpdf\ catalog and, as I will show in this paper, not in 
general mathematically valid.
Stacking is nonetheless considered the preferred method for obtaining \Nz\ from 
a \pzpdf\ catalog \citep{sheldon_photometric_2012, Kelly2014, Benjamin2013, 
bonnett_using_2015, Viironen2015, Asorey2016}.  
However, it must be noted here that Equation~\ref{intro:eq:eqn:stack} is not in 
general mathematically valid.  
(See \citet{Hogg2012} for a complete discussion.)

@@

The redshift distribution function \nz\ is the number of galaxies per unit 
redshift, effectively defining the evolution in the number of galaxies 
\citep{Menard2013}.  
In the following sections, I present and compare methods for estimating \nz\ 
from \pzpdf s.  
\Sect{sec:prob} contains the mathematical derivation of a probabilistic model 
for $n(z)$ dependent on photo-$z$ probability distribution functions, and 
\Sect{sec:sheldon} contrasts the probabilistic model with alternative methods.

@@

We begin by reframing the redshift distribution \nz\ from a probabilistic 
perspective.
Here we define a redshift density \nz\ as the normalized probability density
\begin{equation}
\label{eqn:nz}
\int_{-\infty}^{\infty}\ n(z)\ dz\ \equiv\ \frac{1}{J}\ 
\int_{-\infty}^{\infty}\ \sum_{j=1}^{J}\ \delta(z_{j},\ z)\ dz = 1
\end{equation}
of finding a galaxy $j$ in a catalog of $J$ galaxies having a redshift $z$.
We believe that galaxy redshifts are indeed drawn from \nz, making it a 
probability density over redshift; this fact can also be confirmed by 
dimensional analysis of \Eq{eqn:nz}, as suggested in \citet{Hogg2012}.

We may without loss of generality impose a parameterization
\begin{equation}
\label{eqn:fz}
f(z; \ndphi)\ \equiv\ n(z)
\end{equation}
in terms of some parameter vector $\ndphi$.
At this point, the parameter vector is quite general and may represent 
coefficients in a high-order polynomial as a function of redshift, a set of 
means and variances defining Gaussians that sum to the desired distribution, a 
set of histogram heights that describe a binned version of the redshift 
distribution function, etc.
Upon doing so, we may rewrite \Eq{eqn:fz} as 
\begin{equation}
\label{eqn:pz}
z_{j}\ \sim\ \pr{z \gvn \ndphi}\ \equiv\ f(z; \ndphi),
\end{equation}
a probability density over redshift conditioned on the parameters $\ndphi$ 
specifying \nz.
Note that $z_{j}$ does not depend on the redshift $z_{j'}$ of some other galaxy 
$j' \neq j$, a statement of the causal independence of galaxy redshifts from 
one another.

@@

However, this problem has additional causal structure that we can acknowledge.
The photometry results from the redshifts, not the other way around.
This is the fundamental assumption upon which \pz\ estimation is based.
The forward model corresponds to first drawing redshifts according to 
\Eq{eqn:pz} and then drawing data from the likelihood
\begin{equation}
\label{eqn:pzpdf}
\data_{j}\ \sim\ \pr{\data \gvn z_{j}}
\end{equation}
of photometry conditioned on redshift, illustrated in 
Figure~\ref{intro:fig:fig:pedagogical_scatter}.

@@

A forward model such as that of \Sect{sec:forward} corresponds to a 
probabilistic graphical model (PGM), represented by a directed acyclic graph 
(DAG) as in \Fig{fig:pgm}.
A DAG conveys the causal relationships between physical parameters and, like a 
Feynman diagram in the context of particle physics, is a shorthand for 
mathematical relationships between variables.
The photometric data $\data_{j}$ of a galaxy is drawn from some function of its 
redshift $z_{j}$, independent of other galaxies' data and redshift.
Both data and redshift are random variables, but data is the one that we 
observe and redshift is not directly observable.
In this problem, we don't care about further constraining the redshifts of 
individual galaxies, only the redshift distribution, so we consider redshift to 
be a \textit{latent variable}.
Because the parameters $\ndphi$ that we seek are causally separated from the 
data by the latent variable of redshift, we call them \textit{hyperparameters}.

\begin{figure}
	\begin{center}
		\includegraphics[width=0.2\textwidth]{figures/chippr/pgm.png}
		\caption{The directed acyclic graph of the CHIPPR model, where 
circles indicate random variables and arrows indicate causal relationships.
			The redshift distribution function parameterized by 
hyperparameters $\ndphi$ exists independent of the survey of $J$ galaxies, 
indicated as a box.  
			The redshifts $\{z_{j}\}$ of all galaxies in the survey 
are latent variables independently drawn from the redshift distribution, which 
is a function of $\ndphi$. 
			The photometric data $\data_{j}$ for each galaxy is 
drawn from a function of its redshift $z_{j}$ and observed, indicated by a 
shaded circle.}
		\label{fig:pgm}
	\end{center}
\end{figure}

@@

The implicit prior $\ndphi^{*}$ may be thought of as an initial guess for the 
parameters contained in $\ndphi$, inspired by the generative model for 
photometry from the redshift distribution functions and including some 
parameters defining intrinsic galaxy spectra and instrumental effects. 
(See \citealt{benitez_bayesian_2000} for more detail.)  
For statistical purposes, we would like any interim prior to be uninformative, 
but this is rarely achievable.  
In the case of estimating \nz\ photometrically, it is common to use 
$\ndphi^{*}$ corresponding to \nz\ derived from some different, 
spectroscopically confirmed sample or from a cosmological simulation.

@@

With these definitions, we obtain the desired expression for $\pr{\ndphi \gvn 
\{\data_{j}\}}$,
\begin{align}
\begin{split}
\label{eqn:fullpost}
\pr{\ndphi \gvn \{\data_{j}\}} & \propto 
\pr{\ndphi}\left[\integral{\prod_{j=1}^{J} \left(\pr{z \gvn \data_{j}, 
\ndphi^{*}}\pr{z \gvn \ndphi}(\pr{z \gvn \ndphi^{*}})^{-1}\right)}{z}\right] ,
\end{split}
\end{align}
the heart of \Chippr.
The derivation is provided below in \Sect{app:math}.

@@

@@

We shall make two assumptions of independence in order to make the problem 
tractable; their limitations are be discussed below.  
First, we take $\ln[\pr{\{\data_{j}\} \gvn \{z_{j}\}}]$ to be the sum of $J$ 
individual log-likelihood distribution functions $\ln[\pr{\data_{j} \gvn 
z_{j}}]$, as in 
\begin{equation}
\label{eqn:indiedat}
\ln[\pr{\{\data_{j}\} \gvn \{z_{j}\}}] = \sum_{j=1}^{J}\ \ln[\pr{\data_{j} \gvn 
z_{j}}],
\end{equation}
a result of the definition of probabilistic independence encoded by the box in 
\Fig{fig:pgm}.
Second, we shall assume the true redshifts $\{z_{j}\}$ are $J$ independent 
draws from the true $\pr{z \gvn \ndphi}$.  
Additionally, $J$ itself is a Poisson random variable.  
The combination of these assumptions is given by 
\begin{equation}
\label{eqn:indie}
\ln[\pr{\{z_{j}\} \gvn \ndphi}] = -\integral{f(z; \ndphi)}{z} + \sum_{j=1}^{J}\ 
\ln[\pr{z_{j} \gvn \ndphi}].
\end{equation}
The derivation differs when $J$ is not known, say, when we want to learn about 
a distribution in nature rather than a distribution specific to data in hand, 
but for a photometric galaxy catalog where the desired quantity is $n(z)$ for 
the galaxies entering a larger cosmology calculation, it is a fixed quantity.
A detailed discussion of this matter may be found in 
\citet{Foreman-Mackey2014}.  
Applying Bayes' Rule, we may combine terms to obtain 
\begin{equation}
\label{eqn:posterior}
\ln[\pr{\ndphi \gvn \{\data_{j}\}}] \propto \ln[\pr{\ndphi}] - \integral{f(z; 
\ndphi)}{z} + \sum_{j=1}^{J}\ln\left[\integral{\pr{\data_{j} \gvn z} \pr{z \gvn 
\ndphi}}{z}\right].
\end{equation}

@@

To perform the necessary transformation from likelihoods to posteriors, we 
follow the reasoning of \citet{Foreman-Mackey2014}.  
Let us consider the probability of the parameters conditioned on the data and 
an interim prior and rewrite the problematic likelihood of \Eq{eqn:posterior} 
as 
\begin{equation}
\label{eqn:trick}
\ln[\pr{\data_{j} \gvn z}] = \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn 
\data_{j}, \ndphi^{*}}] - \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}].
\end{equation}

>>>>

\RequirePackage[]{lineno}

@@

\usepackage{tikz}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{lineno}

@@

\usepackage{soul}
%\usepackage[title]{appendix}

\newcommand{\todo}[3]{{\color{#2}\emph{#1}: #3}}
\newcommand{\aim}[1]{\todo{AIM}{red}{#1}}
\newcommand{\myemail}{aimalz@nyu.edu}
\newcommand{\mathul}[1]{\underline{#1}}
\newcommand{\Sect}[1]{Section~\ref{#1}}
\newcommand{\Eq}[1]{Equation~\ref{#1}}
\newcommand{\Fig}[1]{Figure~\ref{#1}}
\newcommand{\Tab}[1]{Table~\ref{#1}}
\newcommand{\project}[1]{{\textsc{#1}}~}
\newcommand{\lsst}{\project{LSST}}
\newcommand{\desc}{\lsst-\project{DESC}}
\newcommand{\sdss}{\project{SDSS}}
\newcommand{\boss}{\project{BOSS}}
\newcommand{\des}{\project{DES}}
\newcommand{\Chippr}{\project{CHIPPR}}
\newcommand{\repo}[1]{{\texttt{#1}}~}
\newcommand{\qp}{\repo{qp}}
\newcommand{\chippr}{\repo{chippr}}
\newcommand{\cosmolike}{\repo{CosmoLike}}
\newcommand{\emcee}{\repo{emcee}}
\newcommand{\github}{\href{https://github.com}{GitHub}}
\newcommand{\python}{\textit{Python}}
\newcommand{\nz}{$n(z)$}
\newcommand{\data}{\ensuremath{\vec{d}}}
% could change to bold
\newcommand{\like}{\mathscr{L}}
\newcommand{\pr}[1]{\ensuremath{\mathrm{p}(#1)}}
% could change to Prob or Pr 
\newcommand{\expect}[1]{\left<#1\right>}
\newcommand{\normal}[2]{\mathcal{N} (#1, #2)}
\newcommand{\gvn}{\mid}% could use | or \vert
\newcommand{\integral}[2]{\ensuremath{\int\ #1\ \mathrm{d} #2}}
\newcommand{\sz}{spec-$z$~}
\newcommand{\Sz}{Spec-$z$~}
\newcommand{\pz}{photo-$z$~}
\newcommand{\Pz}{Photo-$z$~}
\newcommand{\pzpdf}{\pz PDF}
% could change to posterior
\newcommand{\Pzpdf}{\Pz PDF}
% could change to posterior
\newcommand{\zpdf}{redshift posterior}
% could change to implicit posterior
\newcommand{\pzip}{\pz implicit posterior}
\newcommand{\Nz}{$N(z)$~}
\newcommand{\stack}{$\hat{N}(z)$~}
\newcommand{\ntot}{\ensuremath{N_{\mathrm{tot}}}}
\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
% could change to \vec
\newcommand{\ndphi}{\bvec{\phi}}
\newcommand{\mmle}{marginalized maximum likelihood estimate}
% really marginalized maximum a posteriori estimate, may still change this

\begin{document}
\linenumbers

@@

\author{Alex I. Malz\altaffilmark{1}}
\author{David W. Hogg\altaffilmark{1,2,3,4}}
\email{aimalz@nyu.edu}

@@

\begin{abstract}
A trustworthy estimate of the redshift distribution \nz\ is crucial for using weak gravitational lensing and large-scale structure of galaxy catalogs to study cosmology.
Spectroscopic redshifts for the dim and numerous galaxies of weak-lensing surveys are expected to be unavailable, making photometric redshifts (photo-$z$s) the next-best alternative.
The nontrivial systematics affecting photo-$z$ estimation have motivated the weak-lensing community to favor photo-$z$ probability density functions (PDFs) as a more comprehensive alternative to photo-$z$ point estimates.
However, analytic methods for utilizing these new data products in cosmological inference are still evolving.
The ubiquitous methodology known as stacking produces a systematically biased estimator of $n(z)$ that worsens with decreasing signal-to-noise, the very regime where photo-$z$ PDFs are most necessary.
We introduce a mathematically rigorous probabilistic graphical model (PGM) of $n(z)$ for hierarchical inference, which is provably the only self-consistent way to combine photo-$z$ PDFs to produce an estimator of $n(z)$.
Our Cosmological Hierarchical Inference with Probabilistic Photometric Redshifts (\Chippr) model yields a more accurate characterization of $n(z)$ by correctly propagating the redshift uncertainty information beyond the best-fit estimator produced by traditional procedures.
The CHIPPR approach is applicable to any one-point statistic of any random variable.
We conclude by propagating these effects to constraints in the space of cosmological parameters, .
% be clear that rigor comes at a cost, chippr makes explicit the requirements that the pzpdfs be posteriors under a known prior (which would be required by any justifiable method), be honest that we're not ready
\end{abstract}

@@

\aim{Some of the figures refer to left/right but are shown as top/bottom \textemdash I'm still working through a LaTeX bug.}
\aim{The language is very thesis-centric (``me, me, me''), but I'll change it to the standard ``we'' style soon.}

@@

\aim{TODO: Pull in content from global intro of thesis.}

Though their potential to improve estimates of physical parameters is tremendous, \pzpdf s have been applied only to a limited extent.  
They have been used to form selection criteria of samples from galaxy surveys without propagation 
through the calculations of physical parameters \citep{van_breukelen_reliable_2009, viironen_high_2015}.  
Probability cuts on Bayesian quantities are not uncommon \citep{leung_bayesian_2017, dipompeo_quasar_2015}, but that procedure does not fully take advantage of all information contained in a probability distribution for parameter inference.

@@

The \textit{redshift distribution function \Nz}, or, almost interchangably, its normalized cousin the \textit{redshift density function \nz}, serves as an ideal statistic upon which to demonstrate this novel approach.  
\Nz\ is necessary for calculations of two-point correlation functions of weak gravitational lensing and counting statistics that are used to probe dark energy \citep{masters_mapping_2015}.  
The observed \Nz\ also been used to validate survey selection functions usnfenjed in generation of realistic, multi-purpose mock catalogs \citep{norberg_2df_2002}.  
Additionally, \Nz\ has been the subject of inference using \pzpdf s before \citep{sheldon_photometric_2012, hildebrandt_cfhtlens:_2012, kelly_weighing_2014, benjamin_cfhtlens_2013, bonnett_using_2015, viironen_high_2015, asorey_galaxy_2016, leistedt_hierarchical_2016}, so comparisons to the literature may easily be made.

The prevailing way to estimate \nz\ for a sample of $\ntot$ galaxies $i$ is to ``stack'' \pzpdf s according to \aim{fix reference}.
% Equation~\ref{intro:eq:eqn:stack}.
The stacked estimator of the redshift density function is effectively an average of the \pzpdf\ catalog and, as I will show in this paper, not in general mathematically valid.
Stacking is nonetheless considered the preferred method for obtaining \Nz\ from a \pzpdf\ catalog \citep{sheldon_photometric_2012, kelly_weighing_2014, benjamin_cfhtlens_2013, bonnett_using_2015, viironen_high_2015, asorey_galaxy_2016}.  
However, it must be noted here that \aim{fix reference}
%Equation~\ref{intro:eq:eqn:stack} 
is not in general mathematically valid.  
(See \citet{hogg_data_2012} for a complete discussion.)

%\aim{Motivate how well we need to know \Nz\ and why it matters.
%	New paragraph here: Say what precision is needed for \Nz\ for future weak lensing surveys. 
%	Say what precision the mass function is needed (in, say cluster studies) for precision cosmology.}

@@

The redshift distribution function \nz\ is the number of galaxies per unit redshift, effectively defining the evolution in the number of galaxies \citep{menard_clustering-based_2013}.  
% \aim{also selection function}
In the following sections, I present and compare methods for estimating \nz\ from \pzpdf s.  
\Sect{sec:prob} contains the mathematical derivation of a probabilistic model for $n(z)$ dependent on photo-$z$ probability distribution functions, and \Sect{sec:sheldon} contrasts the probabilistic model with alternative methods.

@@

We begin by reframing the redshift distribution \nz\ from a probabilistic perspective.
Here we define a redshift density \nz\ as the normalized probability density
\begin{equation}
\label{eqn:nz}
\int_{-\infty}^{\infty}\ n(z)\ dz\ \equiv\ \frac{1}{J}\ \int_{-\infty}^{\infty}\ \sum_{j=1}^{J}\ \delta(z_{j},\ z)\ dz = 1
\end{equation}
of finding a galaxy $j$ in a catalog of $J$ galaxies having a redshift $z$.
We believe that galaxy redshifts are indeed drawn from \nz, making it a probability density over redshift; this fact can also be confirmed by dimensional analysis of \Eq{eqn:nz}, as suggested in \citet{hogg_data_2012}.

We may without loss of generality impose a parameterization
\begin{equation}
\label{eqn:fz}
f(z; \ndphi)\ \equiv\ n(z)
\end{equation}
in terms of some parameter vector $\ndphi$.
At this point, the parameter vector is quite general and may represent coefficients in a high-order polynomial as a function of redshift, a set of means and variances defining Gaussians that sum to the desired distribution, a set of histogram heights that describe a binned version of the redshift distribution function, etc.
Upon doing so, we may rewrite \Eq{eqn:fz} as 
\begin{equation}
\label{eqn:pz}
z_{j}\ \sim\ \pr{z \gvn \ndphi}\ \equiv\ f(z; \ndphi),
\end{equation}
% \aim{explain "is drawn from"}
a probability density over redshift conditioned on the parameters $\ndphi$ specifying \nz.
Note that $z_{j}$ does not depend on the redshift $z_{j'}$ of some other galaxy $j' \neq j$, a statement of the causal independence of galaxy redshifts from one another.

@@

However, this problem has additional causal structure that we can acknowledge.
The photometry results from the redshifts, not the other way around.
This is the fundamental assumption upon which \pz\ estimation is based.
The forward model corresponds to first drawing redshifts according to \Eq{eqn:pz} and then drawing data from the likelihood
\begin{equation}
\label{eqn:pzpdf}
\data_{j}\ \sim\ \pr{\data \gvn z_{j}}
\end{equation}
of photometry conditioned on redshift, illustrated in \aim{fix reference}.
%Figure~\ref{intro:fig:fig:pedagogical_scatter}.

@@

A forward model such as that of \Sect{sec:forward} corresponds to a probabilistic graphical model (PGM), represented by a directed acyclic graph (DAG) as in \Fig{fig:pgm}.
A DAG conveys the causal relationships between physical parameters and, like a Feynman diagram in the context of particle physics, is a shorthand for mathematical relationships between variables.
The photometric data $\data_{j}$ of a galaxy is drawn from some function of its redshift $z_{j}$, independent of other galaxies' data and redshift.
Both data and redshift are random variables, but data is the one that we observe and redshift is not directly observable.
In this problem, we don't care about further constraining the redshifts of individual galaxies, only the redshift distribution, so we consider redshift to be a \textit{latent variable}.
Because the parameters $\ndphi$ that we seek are causally separated from the data by the latent variable of redshift, we call them \textit{hyperparameters}.
%\aim{I'm not explaining the distinction between hyperparameters and parameters adequately.}

\begin{figure}
	\begin{center}
		\includegraphics[width=0.15\textwidth]{figures/chippr/pgm.png}
		\caption{The directed acyclic graph of the CHIPPR model, where circles indicate random variables and arrows indicate causal relationships.
			The redshift distribution function parameterized by hyperparameters $\ndphi$ exists independent of the survey of $J$ galaxies, indicated as a box.  
			The redshifts $\{z_{j}\}$ of all galaxies in the survey are latent variables independently drawn from the redshift distribution, which is a function of $\ndphi$. 
			The photometric data $\data_{j}$ for each galaxy is drawn from a function of its redshift $z_{j}$ and observed, indicated by a shaded circle.}
		\label{fig:pgm}
	\end{center}
\end{figure}

@@

%Posteriors differ from likelihoods by way of a prior distribution, so we cannot simply assume that the available data products are \pz\ posteriors $\pr{z \gvn \data_{j}}$.  
%Rather, we have a catalog of implicit-prior weighted \pz\ posteriors $\pr{z \gvn \data_{j}, \ndphi^{*}}$.  
%There must have been some interim prior probability distribution $p(z|\vec{\theta}^{*})$ defined in terms of the interim prior parameter values (hereafter the interim prior) $\vec{\theta}^{*}$ explicitly chosen or implicitly made to perform the calculation of the probabilistic photo-$z$s.  
%If it is implicit, it may not be representable in the parametrization we have chosen, and furthermore it may not be known at all; a method that produces interim photo-$z$ posteriors of this kind is not suitable for inference.  
%However, so long as the implicit prior is known, hierarchical inference is possible.

The implicit prior $\ndphi^{*}$ may be thought of as an initial guess for the parameters contained in $\ndphi$, inspired by the generative model for photometry from the redshift distribution functions and including some parameters defining intrinsic galaxy spectra and instrumental effects. 
(See \citet{benitez_bayesian_2000} for more detail.)  
For statistical purposes, we would like any interim prior to be uninformative, but this is rarely achievable.  
In the case of estimating \nz\ photometrically, it is common to use $\ndphi^{*}$ corresponding to \nz\ derived from some different, spectroscopically confirmed sample or from a cosmological simulation.

@@

% this is what we want to compute
% what we're given is implicit posteriors
% this turns out to be the correct answer
% derivation in appendix
% human-readable description of ratios replaces the implict prior with model at each sample step, converting posterior into likelihood
With these definitions, we obtain the desired expression for $\pr{\ndphi \gvn \{\data_{j}\}}$,
\begin{align}
\begin{split}
\label{eqn:fullpost}
\pr{\ndphi \gvn \{\data_{j}\}} \propto & \pr{\ndphi} \left[\integral{\prod_{j=1}^{J} \right.\\
	& \left. \left(\pr{z \gvn \data_{j}, \ndphi^{*}} \pr{z \gvn \ndphi}(\pr{z \gvn \ndphi^{*}})^{-1}\right)}{z}\right] ,
\end{split}
\end{align}
the heart of \Chippr.
The derivation is provided below in \Sect{app:math}.

@@

\aim{This was originally envisioned as an appendix, and my default plan is to move it back to one.}

%We begin by parametrizing $N(z)$ in terms of $\vec{\theta}$, comprising some set of hyperparameters that define the form $N(z)$ may take in whatever basis we choose.  
%We define a function $f_{\vec{\theta}}(z)=N(z)$ that transforms these hyperparameters into the redshift distribution function $N(z)$.  
%Because 
%\begin{equation}
%\eqlabel{eq:definition}
%N(z) \propto p(z \gvn \vec{\theta}),
%\end{equation}
%we may discontinue discussion of $N(z)$ in favor of the likelihood $p(z|\vec{\theta})$.

@@

We shall make two assumptions of independence in order to make the problem tractable; their limitations are be discussed below.  
First, we take $\ln[\pr{\{\data_{j}\} \gvn \{z_{j}\}}]$ to be the sum of $J$ individual log-likelihood distribution functions $\ln[\pr{\data_{j} \gvn z_{j}}]$, as in 
\begin{equation}
\label{eqn:indiedat}
\ln[\pr{\{\data_{j}\} \gvn \{z_{j}\}}] = \sum_{j=1}^{J}\ \ln[\pr{\data_{j} \gvn z_{j}}],
\end{equation}
a result of the definition of probabilistic independence encoded by the box in \Fig{fig:pgm}.
Second, we shall assume the true redshifts $\{z_{j}\}$ are $J$ independent draws from the true $\pr{z \gvn \ndphi}$.  
Additionally, $J$ itself is a Poisson random variable.  
The combination of these assumptions is given by 
\begin{equation}
\label{eqn:indie}
\ln[\pr{\{z_{j}\} \gvn \ndphi}] = -\integral{f(z; \ndphi)}{z} + \sum_{j=1}^{J}\ \ln[\pr{z_{j} \gvn \ndphi}].
\end{equation}
%It is important to note that the integral $\integral{n(z)}{z} N(z)\ dz$ is not constrained to equal the variable defining the Poisson distribution but instead $J$ by \Eq{eq:definition}, which can be thought of as another parameter.  
The derivation differs when $J$ is not known, say, when we want to learn about a distribution in nature rather than a distribution specific to data in hand, but for a photometric galaxy catalog where the desired quantity is $n(z)$ for the galaxies entering a larger cosmology calculation, it is a fixed quantity.
A detailed discussion of this matter may be found in \citet{foreman-mackey_exoplanet_2014}.  
Applying Bayes' Rule, we may combine terms to obtain 
\begin{align}
\begin{split}
\label{eqn:posterior}
\ln[\pr{\ndphi \gvn \{\data_{j}\}}] & \propto \ln[\pr{\ndphi}] - \integral{f(z; \ndphi)}{z}\\
& + \sum_{j=1}^{J}\ln\left[\integral{\pr{\data_{j} \gvn z} \pr{z \gvn \ndphi}}{z}\right].
\end{split}
\end{align}

%\Eq{eq:posterior} contains two quantities that merit further discussion, the prior distribution $p(\vec{\theta})$ discussed further in \Sect{sec:exp} and the photo-$z$ log-likelihoods $\ln[p(\vec{d}_{j}|z_{j})]$ that have not been mentioned since \Eq{eq:marginalize}.  
%Though photo-$z$ log-likelihoods would be desirable for use in these equations, they are not generally the product of either empirical and data-driven methods for obtaining photo-$z$ probability distributions.  
%Though probabilistic photo-$z$s are typically reported as generic probability distributions $p(z_{j})$, the methods that produce them may be understood to always yield posteriors, probability distributions conditioned on the data we believe to be true.  
%If they were not based in this assumption, they would require a sum over an infinite space of possible datasets.

@@

To perform the necessary transformation from likelihoods to posteriors, we follow the reasoning of \citet{foreman-mackey_exoplanet_2014}.  
Let us consider the probability of the parameters conditioned on the data and an interim prior and rewrite the problematic likelihood of \Eq{eqn:posterior} as 
\begin{equation}
\label{eqn:trick}
\ln[\pr{\data_{j} \gvn z}] = \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] - \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}].
\end{equation}

Once the implicit prior $\ndphi^{*}$ is explicitly introduced, we may expand the last term in \Eq{eqn:trick} according to Bayes' Rule to get 
\begin{align}
\begin{split}
\label{eqn:expand}
\ln[\pr{\data_{j} \gvn z}] & = \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] + \ln[\pr{\data_{j} \gvn \ndphi^{*}}]\\
& - \ln[\pr{z \gvn \ndphi^{*}}] - \ln[\pr{\data_{j} \gvn z, \ndphi^{*}}].
\end{split}
\end{align}
Because there is no direct dependence of the data upon the hyperparameters, we may again expand the term $\ln[\pr{\data_{j} \gvn z, \ndphi^{*}}]$ to obtain 
\begin{align}
\begin{split}
\label{eqn:indterm}
\ln[\pr{\vec{d}_{j} \gvn z}] & = \ln[\pr{\data_{j} \gvn z}] + \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] + \ln[\pr{\data_{j} \gvn \ndphi^{*}}] \\
& - \ln[\pr{z \gvn \ndphi^{*}}] - \ln[\pr{\data_{j} \gvn z}] - \ln[\pr{\data_{j} \gvn \ndphi^{*}}].
\end{split}
\end{align}
Canceling the undesirable terms for the inaccessible likelihood $\ln[\pr{\data_{j} \gvn z}]$ and trivial $\ln[\pr{\data_{j} \gvn \ndphi^{*}}]$ yields
\begin{equation}
\label{eqn:cancel}
\ln[\pr{\data_{j} \gvn z}] = \ln[\pr{z \gvn \data_{j}, \ndphi^{*}}]  - \ln[\pr{z \gvn \ndphi^{*}}].
\end{equation}
We put this all together to get the full log-posterior probability distribution of 
\begin{align}
\begin{split}
\label{eqn:final}
\ln[\pr{\ndphi \gvn \{\data_{j}\}}] & \propto \ln[\pr{\ndphi}] + \ln \left[\integral{\exp \left[\sum_{j=1}^{J} \right.\right.\\
& \left.\left.\left(\ln[\pr{z \gvn \data_{j}, \ndphi^{*}}] + \ln[\pr{z \gvn \ndphi}] - \ln[\pr{z \gvn \ndphi^{*}}]\right)\right]}{z}\right] ,
\end{split}
\end{align}
